On Optimality of Finite Element Integration

### 2016-02-16 TOMS_Paper/assembly.tex
Swith to ILP formulation (TBF)

\item Use rewrite rules to restructure the expression. Let $V = \lbrace v^1, ..., v^n \rbrace$ be the vertices adjacent to $s \in S$; the preconditions for the three rewrite rules (A), (B), and (C) are:
\begin{enumerate}[label=(\Alph*)]
\item $A \subsetneq V$ is a set of vertices not standing on cyclic paths, $A \neq \emptyset$
\item $s$ is in a cycle $C$ and has highest degree among all vertices in $C$
\item $\lbrace v^1, ..., v^{n-1} \rbrace$ are leaves of a tree, $p = v^n$ is the parent of $s$ (unless $s$ is the root)
\end{enumerate}
If a precondition is met, we take $s$ and apply Strategy~\ref{strategy:ii} (i.e., we factorize $s$ and perform code motion), and we remove all edges incident to $s$. The preconditions are checked cyclically (first all (A)s, then all (B)s, with (B) that may expose further hits for (A), and finally all (C)s) until no more hits are possible (i.e., until $S = T$).  \\
\textit{Note: steps (4) and (5) only impact bilinear forms, since otherwise $E = \emptyset$.}

@@

\subsubsection{Examples}
\label{sec:se-examples}
Consider again Figure~\ref{code:multi_loopnest_a}. We have $\mathbb{P} = \lbrace P^0, P^1, P^2\rbrace$, with $P^0 = \lbrace b_j\rbrace$, $P^1 = \lbrace c_i\rbrace$, and $P^2 = \lbrace d_i\rbrace$. For all $P^i$, we have $|P^i| = 1 = |S_{P^i}|$, although applying Strategy~\ref{strategy:i} in step (1) has no effect. The sharing graph is $G = (\lbrace b_j, c_i, d_i \rbrace, \lbrace (b_j, c_i), (b_j, d_i) \rbrace$, and $T = \lbrace c_i, d_i \rbrace$. Rewrite rule (4A) is hit, which leads to the code in Figure~\ref{code:multi_loopnest_c}.

In Figure~\ref{code:poisson}, Algorithm~\ref{algo:sharing-elimination} is executed in a realistic scenario, which originates from the bilinear form of a Poisson equation in two dimensions. We observe that $\mathbb{P} = \lbrace P^0, P^1\rbrace$, with $P^0 = \lbrace (z_0 a_{ik} + z_2 b_{ik}), (z_1 a_{ik} + z_3 b_{ik})\rbrace$ and $P^1 = \lbrace (z_0 a_{ij} + z_2 b_{ij}), (z_1 a_{ij} + z_3 b_{ij})\rbrace$. In addition, $|P^i| = |S_{P^i}| = 2$, so Strategy~\ref{strategy:i} is applied to both partitions (step (1)). We then have (step (2)) $G = (\lbrace t^0, t^1, t^2, t^3 \rbrace, \lbrace (t^0, t^2), (t^1, t^3) \rbrace)$, although no rewrite rule (step (5)) is applicable since $S = T$

@@

After structuring the sharing graph $G$ in such a way that only flop-decreasing transformations are possible, a set of rewrite rules are applied. We here discuss why applying in order rules (A), (B) and (C) leads to the optimum.

Rule (A) is the critical one because it coordinates factorization at the junction between a cyclic and an acyclic components. It exploits the observation that a vertex outside of cyclic paths may be ``aggregated'' within one or more ``potentially unavoidable'' factorizations, thus saving on the final number of multiplications. This is quite evident for terminals (for which there is at most one way of being factorized), although not so obvious for vertices of arbitrary degree. Let $a$ be a vertex outside of cyclic paths, and let $V = \lbrace v^1, ..., v^n \rbrace$ be the set of vertices adjacent to $a$. Consider the non-trivial case in which at least one vertex $s \in V$ is within a cycle. If $a$ and $s$ were factorized separately, we could end up with the two sub-expressions $a(...v^1 + ... s + ...v^n)$ and $s(...)$. The spatial sharing induced by $s$ would make the synthesis sub-optimal. Conversely, if $s$ is factorized such that a single sub-expression $s(...a + ...)$ will be produced, the operation count is at worst unchanged.

Rule (B) acts within cycles by iteratively picking the vertex with highest degree. Since the assumption is that there are no more hits for rule (A), the problem here is as simple as selecting the minimum number of vertices whose incident edges would cover the whole cycle. With a very similar concept, rule (C) schedules the remaining (disconnected) sub-trees.
\end{enumerate}
\end{proof}

>>>>

\item Map $G$ to an Integer Linear Programming (ILP) problem for determining how to optimally apply Strategy~\ref{strategy:ii}. The solution is the set of symbols that will be factorized by Strategy~\ref{strategy:ii}. Let $|S| = n$; the ILP model then is as follows:
\begin{gather*}
x_i \text{: a vertex in } S \text{ (1 if a symbol must be factorized, 0 otherwise)}\\
y_{ij} \text{: a direct edge in } E \text{ (1 if the product } s^i s^j \text{ is factorized through } s^i \text{, 0 otherwise)}\\
n_i \text{: the number of edges incident to } x_i \\
\begin{align*}
\min \sum_{i=1}^{n} x_i,\ s.t. ~~~~~~~~~&\sum_{j=1}^{n} y_{ij} \leq n_i x_i,\ i = 1, ..., n \\
&\sum_{i=1}^{n} y_{ij} = 1,\ j = 1, ..., n
\end{align*}
\phantom{\hspace{6cm}}
\end{gather*}

@@

\subsubsection{Examples}
\label{sec:se-examples}
Consider again Figure~\ref{code:multi_loopnest_a}. We have $\mathbb{P} = \lbrace P^0, P^1, P^2\rbrace$, with $P^0 = \lbrace b_j\rbrace$, $P^1 = \lbrace c_i\rbrace$, and $P^2 = \lbrace d_i\rbrace$. For all $P^i$, we have $|P^i| = 1 = |S_{P^i}|$, although applying Strategy~\ref{strategy:i} in step (1) has no effect. The sharing graph is $G = (\lbrace b_j, c_i, d_i \rbrace, \lbrace (b_j, c_i), (b_j, d_i) \rbrace$, and $T = \lbrace c_i, d_i \rbrace$. The ILP formulation leads to the code in Figure~\ref{code:multi_loopnest_c}.

In Figure~\ref{code:poisson}, Algorithm~\ref{algo:sharing-elimination} is executed in a realistic scenario, which originates from the bilinear form of a Poisson equation in two dimensions. We observe that $\mathbb{P} = \lbrace P^0, P^1\rbrace$, with $P^0 = \lbrace (z_0 a_{ik} + z_2 b_{ik}), (z_1 a_{ik} + z_3 b_{ik})\rbrace$ and $P^1 = \lbrace (z_0 a_{ij} + z_2 b_{ij}), (z_1 a_{ij} + z_3 b_{ij})\rbrace$. In addition, $|P^i| = |S_{P^i}| = 2$, so Strategy~\ref{strategy:i} is applied to both partitions (step (1)). We then have (step (2)) $G = (\lbrace t^0, t^1, t^2, t^3 \rbrace, \lbrace (t^0, t^2), (t^1, t^3) \rbrace)$, and since $S = T$ the ILP formulation becomes irrelevant because there are no factorization opportunities.

@@

After structuring the sharing graph $G$ in such a way that only flop-decreasing transformations are possible, the ILP model is instantiated. At this point, the optimality of the algorithm reduces to prove the correctness of the model. The model aims ...
\end{enumerate}
\end{proof}
